---
title: "Simulation-based analysis of animal-vessel interactions with intersimR"
subtitle: "REDUCE Workshop 2026"
author: 
  - name: Dr. David March
    orcid: 0000-0002-6118-761X
    email: david.march@uv.es
    affiliation: Spatial Marine Lab, University of Valencia
date: 2026-05-27
execute:
  eval: true
  echo: true
  warning: false
  message: false
format:
  html:
    toc: true
    toc-depth: 3
---

# Course objectives

This practical session introduces a simulation-based framework to quantify and interpret animal–vessel interactions using high-resolution animal tracking data and vessel trajectory data. The session is centred on the `intersimR` package, which implements tools to identify proximity events, simulate realistic alternative animal trajectories, and test whether observed interactions are more consistent with directed behavioural responses than with random co-occurrence.

Participants are assumed to be familiar with the preprocessing and handling of GPS tracking data in R, including temporal formatting, coordinate systems, and basic trajectory visualization.

By the end of the session, participants should be able to:

-   understand the conceptual distinction between encounters, associations, attraction, and following behaviour;
-   identify proximity-based interaction events using distance and duration thresholds;
-   run simulation-based tests for attraction and following behaviour;
-   interpret the resulting p-values, classifications, and diagnostic plots;
-   understand how the workflow can be scaled from a single animal–vessel pair to larger datasets involving multiple individuals and vessels.

The session prioritizes a simple, reproducible worked example. Rather than aiming to cover all possible analytical configurations, the focus is on understanding the logic of the method and the structure of the workflow. More advanced applications, including batch processing across multiple animals and vessels, sensitivity analyses, and integration into larger ecological studies, are introduced conceptually through the accompanying slides.

# Setup

## Install intersimR

The `intersimR` package is currently installed from GitHub. GitHub installation requires the `remotes` package, which allows R to install packages that are not yet available from CRAN.

```{r}
#| eval: false

install.packages("remotes")
```

Once `remotes` is installed, `intersimR` can be installed using:

```{r}
#| eval: false

remotes::install_github("SpatialMarine/intersimR")
```

If installation fails with `HTTP error 401: Bad credentials`, R is probably using an invalid GitHub token. Run:

```{r}
#| eval: false

Sys.unsetenv("GITHUB_TOKEN")
remotes::install_github("SpatialMarine/intersimR")
```

If the package is updated during the course, it can be reinstalled using the same command.

You can also check the installation and the version of `intersimR` using:

```{r}
packageVersion("intersimR")
```

## Loading packages

Load the packages required for the session. In addition to `intersimR`, we will use:

- `terra` and `sf` for spatial data handling;
- `ggplot2` for visualization;
- `rnaturalearth` for land polygon data;
- `tidyterra` for plotting raster objects with `ggplot2`;
- `gganimate` and `gifsky` for animated trajectory visualizations;
- `readr` for importing example datasets.


```{r}
# Load packages
library(intersimR)
library(terra)
library(sf)
library(ggplot2)
library(rnaturalearth)
library(gganimate)
library(tidyterra)
library(gifski)
library(readr)
```

# Import and visualize example tracking data

In this section we load a simplified animal–vessel tracking dataset. The data have already been pre-processed and formatted, so we will focus on visualizing the tracks and preparing them for the interaction analysis.

```{r}
# Load example data
bird <- readr::read_csv("data/608_bird.csv", show_col_types = FALSE)
ship <- readr::read_csv("data/608_ship.csv", show_col_types = FALSE)
```

## Convert data to `track_tbl` objects

The `intersimR` workflow uses a common track structure for both animal and vessel trajectories. This structure requires longitude, latitude, time, and an individual identifier.

```{r}
bird_trk <- as_track(
  bird,
  lon = longitude,
  lat = latitude,
  time = time,
  id = organismID
)

ship_trk <- as_track(
  ship,
  lon = longitude,
  lat = latitude,
  time = time,
  id = shipID
)
```

## Inspect the data

```{r}
dplyr::glimpse(bird_trk)
dplyr::glimpse(ship_trk)
```

## Static visualization of animal and vessel tracks

A first inspection of the trajectories helps to identify whether the animal and vessel occupied similar areas during the same time window.

```{r}
p <- ggplot() +
  # plot vessel track
  geom_path(
    data = ship_trk,
    aes(x = lon, y = lat, group = id, colour = "Vessel"),
    linewidth = 0.6
  ) +
  geom_point(
    data=ship_trk,
    aes(x=lon, y=lat, group = id, colour="Vessel"),
    size=2) +
  # plot animal track
  geom_path(
    data = bird_trk,
    aes(x = lon, y = lat, group = id, colour = "Animal"),
    linewidth = 0.6
  ) +
  geom_point(
    data=bird_trk,
    aes(x=lon, y=lat, group = id, colour="Animal"),
    size=2) +
  # set colours
  scale_colour_manual(
    values = c(
      "Animal" = "#0072B2",
      "Vessel" = "#E69F00"
    ),
    name = NULL
  ) +
  # aesthetics
  coord_equal() +
  labs(
    x = "Longitude",
    y = "Latitude"
  ) +
  theme_bw()

print(p)
```

## Animated visualization

The static map shows the spatial overlap between trajectories, but animal–vessel interactions are inherently spatio-temporal. An animation can help visualize whether the animal and vessel were close to each other at the same time.

```{r}
p_anim <- p +
  labs(
    x = "Longitude",
    y = "Latitude",
    title = "Time: {frame_along}"
  ) +
  gganimate::transition_reveal(along = time)

gganimate::animate(
  p_anim,
  duration = 6,
  fps = 5,
  renderer = gganimate::gifski_renderer()  # alternative: av_renderer()
)
```

# Matching animal and vessel positions

Before identifying interaction events, animal and vessel locations need to be paired in time. The function `match_animal_vessels()` interpolates or matches vessel positions to the animal time steps, retaining only animal–vessel pairs within a broad spatial and temporal search window.

This step is particularly important when working with large tracking datasets, as it substantially reduces data volume prior to the interaction analysis. There is no need to evaluate animal–vessel pairs that are clearly too far apart to realistically interact. By filtering vessel positions within a broad search radius, the workflow focuses computational effort only on potentially relevant animal–vessel pairs.

Before matching animal and vessel positions, we first assign a unique identifier to every animal location using `add_fix_id()`. This creates an internal fix identifier that allows the workflow to track individual positions throughout the different processing steps and facilitates linking outputs back to the original trajectory:

```{r}
bird_trk <- add_fix_id(bird_trk)
```

The distance threshold should be large enough to capture all potential interactions, but not so large as to include many irrelevant pairs. The temporal tolerance accounts for small time mismatches between animal and vessel positions, which can arise from differences in tracking resolution or interpolation. In this example, we use a broad distance threshold of 30 km and a temporal tolerance of 5 minutes.

```{r}
matched <- match_animal_vessels(
  animal = bird_trk,
  vessel = ship_trk,
  dist_thr_m = 30000,
  time_thr_min = 5
)
```

Exploring the resulting object shows that the original animal track has been filtered to retain only positions that had at least one vessel position within the specified distance and time thresholds. The resulting `matched` object contains paired animal and vessel positions that can be used to detect proximity events.

```{r}
dplyr::glimpse(matched)
```

# Detecting proximity events

The same function, `detect_proximity_events()`, can be used to identify different types of proximity events depending on the distance and duration thresholds selected.

In this workflow, we distinguish between:

-   **encounters**, defined as broad-scale proximity events;
-   **associations**, defined as close and sustained proximity events.

These definitions are analytical choices and should be adapted to the study system, species, vessel type, tracking resolution, and ecological question.

## Detect encounters

Encounters represent broad-scale situations in which an animal and a vessel are close enough to potentially interact or perceive each other.

Here, we define encounters as events in which the animal and vessel are within 30 km for at least 30 minutes. Consecutive events separated by gaps shorter than 60 minutes are merged.

```{r}
encounter <- detect_proximity_events(
  pairs = matched,
  min_dist_m = 30000,
  min_duration_min = 30,
  max_gap_min = 60
)
```

The function `detect_proximity_events()` returns a list containing two main objects:

-   `events`: a summary table describing the detected proximity events;
-   `data`: the subset of paired animal–vessel positions belonging to those events.

The `events` object contains event-level information such as: - event identifier, - start and end time, - event duration, - minimum distance between animal and vessel, - and the number of matched positions included in the event.

The `data` object contains the full time series of paired positions associated with each event identifier and is later used for simulation-based analyses.

```{r}
encounter_events <- encounter$events
encounter_data <- encounter$data

head(encounter_events)
```

## Detect associations

Associations represent close and sustained animal–vessel co-occurrences. These are the candidate events that will later be tested using simulations.

Here, we define associations as events in which the animal and vessel are within 1.5 km for at least 15 minutes. Consecutive associations separated by gaps shorter than 30 minutes are merged.

```{r}
association <- detect_proximity_events(
  pairs = matched,
  min_dist_m = 1500,
  min_duration_min = 15,
  max_gap_min = 30
)
```

```{r}
association_events <- association$events
association_data <- association$data
```

At this stage, it is useful to compare the outputs from encounters and associations.

```{r}
head(encounter_data)
head(association_data)
```

Notice that the eventID column in association_data is only assigned to fixes that satisfy the association criteria (i.e. distance and duration thresholds defined for associations). In contrast, encounter_data includes a broader set of animal–vessel positions corresponding to the less restrictive encounter definition.

This distinction is important because associations represent the subset of proximity events that will later be evaluated using simulation-based tests for attraction and following behaviour.

At this stage, the detected association only indicates close spatio-temporal proximity. It does not yet demonstrate whether the animal was attracted to the vessel or whether it actively followed the vessel. These behavioural interpretations require comparison with simulated animal trajectories, which is addressed in the next section.

# Create an ocean mask

The simulation step generates alternative animal trajectories that represent plausible movement paths under a null model. Because the study system is marine, simulated trajectories should be constrained to remain in the ocean and avoid unrealistic movements over land.

For this reason, we create an ocean mask covering the spatial extent of the animal and vessel tracks.

## Download land polygons

We use land polygons from Natural Earth to define areas that should be excluded from simulated marine trajectories.

```{r}
land_sf <- rnaturalearth::ne_download(
  scale = 10,
  type = "land",
  category = "physical",
  returnclass = "sf"
)
```

## Generate the ocean mask

The function `create_oceanmask()` is a helper function implemented in `intersimR` to facilitate the generation of marine masks for trajectory simulations. The function returns a raster object in which ocean cells are coded as `0` and land cells as `1`. This mask is later used to constrain simulated trajectories to marine areas and avoid unrealistic movements across land.

We will generate a larger mask covering the entire study region and reuse it across all simulations.

The spatial resolution of the mask should be adapted to the study system. Coastal environments with complex shorelines may require finer resolutions, whereas pelagic systems can generally be represented using coarser masks with substantially lower computational cost.

The following code generates a single regional ocean mask for a large portion of the Western Mediterranean Sea:

```{r}
oceanmask <- create_oceanmask(
  bbox = c(-6, 16, 34.5, 45),
  res = 0.005,
  polygon = land_sf,
  polygon_type = "land"
)

plot(oceanmask)
```

## Visualize the mask with the tracking data

```{r}

# define plot boundaries
buffer <- 0.1
xlim_track <- range(c(bird_trk$lon, ship_trk$lon), na.rm = TRUE) + c(-buffer, buffer)
ylim_track <- range(c(bird_trk$lat, ship_trk$lat), na.rm = TRUE) + c(-buffer, buffer)


ggplot() +
  geom_spatraster(
    data = as.factor(oceanmask)
  ) +
  scale_fill_manual(
    values = c("0" = "white", "1" = "grey10"),
    guide = "none") +
  geom_path(
    data = ship_trk,
    aes(x = lon, y = lat, colour = "Vessel"),
    linewidth = 0.6
  ) +
  geom_path(
    data = bird_trk,
    aes(x = lon, y = lat, colour = "Animal"),
    linewidth = 0.6
  ) +
  scale_colour_manual(
    values = c(
      "Animal" = "#0072B2",
      "Vessel" = "#E69F00"
    ),
    name = NULL
  ) +
  coord_sf(
    xlim = xlim_track,
    ylim = ylim_track,
    expand = FALSE
  ) +
  labs(
    x = "Longitude",
    y = "Latitude"
  ) +
  theme_bw() +
  theme(
  panel.grid = element_blank()
)
```

At this point, the data are ready for simulation-based testing. In the next section, we will extract one association event and test whether the observed animal movement is consistent with attraction towards the vessel.

# Attraction analysis

The attraction analysis evaluates whether the animal moved towards the vessel more directly than expected under random movement.

The rationale of the test is illustrated below:

1.  the observed animal trajectory prior to the association is extracted;
2.  alternative trajectories are simulated using the movement characteristics of the observed track;
3.  the observed trajectory is compared against the simulated trajectories;
4.  if the observed animal reaches the vessel more consistently than expected under the simulations, the interaction is classified as significant for attraction.

Importantly, attraction is evaluated *before* the association starts. Therefore, the extracted trajectory segment must include a temporal window preceding the association event.

## Select an association event

First, inspect the detected association events.

```{r}
association_events
```

For this example, we will analyse the first detected association.

```{r}
event_id <- association_events$eventID[1]

event_id
```

## Extract the trajectory segment

The function `extract_event_segments()` extracts the animal and vessel trajectories associated with the selected event.

For the attraction analysis, we include a temporal window preceding the association onset. In this example, we extract the 30 minutes before the association starts.

```{r}
attraction_seg <- extract_event_segments(
  x = association,
  eventID = event_id,
  lead_secs = 1800,
  lag_secs = 0
)
```

The extracted object contains a list of three elements:

-   the animal trajectory segment;
-   the vessel trajectory segment;
-   metadata describing the event.

```{r}
attraction_seg$meta
```


# Running the attraction test

The attraction test compares the observed animal trajectory against simulated trajectories generated under a null movement model.

The simulations preserve key movement characteristics of the observed trajectory, including autocorrelation and directional persistence, while randomizing the spatial path. The observed trajectory is then compared against these simulated alternatives to evaluate whether the animal approached the vessel more directly than expected under random movement.

## Run the simulations

The function `test_interactions()` performs the simulation-based test.

In this example:

-   `method = "attract"` specifies that we are testing attraction behaviour;
-   `sim_n = 100` generates 100 simulated trajectories;
-   `oceanmask` constrains simulations to marine areas;
-   `anchor = "start"` initializes simulations from the observed starting position.
-   `min_locs = 6` defines the minimum number of locations required to fit the movement model and run the simulations;
-   `cores = 2` allows simulations to run in parallel using two processor cores, reducing computation time;
-   `seed = 123` fixes the random number generator seed to ensure reproducibility of the simulations;
-   `return_simdata = TRUE` stores the simulated trajectories in the output object, allowing later visualization and diagnostic analyses.

```{r}
test_attraction <- test_interactions(
  animal = attraction_seg$animal,
  vessel = attraction_seg$vessel,
  obs_duration_min = NULL,
  min_dist_m = 1500,
  min_duration_min = 15,
  max_gap_min = 30,
  method = "attract",
  sim_n = 100,
  oceanmask = oceanmask,
  anchor = "start",
  min_locs = 6,
  cores = 2,
  seed = 123,
  return_simdata = TRUE
)
```

Depending on the number of simulations and trajectory length, this step may take several seconds.

## Inspect the output

```{r}
names(test_attraction)
```

The output contains different listed elements, including:

-   `result`: summary statistics and p-values;
-   `meta`: metadata describing the simulation settings;
-   `sim_pairs`: simulated animal–vessel paired positions;
-   `sim_events`: summary of the proximity events detected in the simulated trajectories.

```{r}
test_attraction$result
```

## Interpreting the attraction test

The attraction test evaluates whether the observed animal trajectory approached the vessel more strongly than expected under the simulated trajectories.

A low p-value (p \< 0.05) indicates that the observed trajectory is unlikely to result from random movement alone and is therefore considered significant for attraction behaviour.

In contrast, non-significant results indicate that the observed approach could plausibly arise from the movement structure of the trajectory itself without requiring directed attraction towards the vessel.

## Visualize simulated trajectories

The simulated trajectories can be compared against the observed trajectory to better understand the test.

```{r}
sim_animal_attract <- as_track(
  test_attraction$sim_pairs,
  lon = animal_lon,
  lat = animal_lat,
  time = animal_time,
  id = matchID
)
```



```{r}
# format p-value for annotation
p_lab <- paste0("p = ", format(round(test_attraction$result$p_value, 3), nsmall = 3,
                                 scientific = FALSE, trim = TRUE))

# define plot boundaries
buffer <- 0.1
xlim_track <- range(c(bird_trk$lon, ship_trk$lon), na.rm = TRUE) + c(-buffer, buffer)
ylim_track <- range(c(bird_trk$lat, ship_trk$lat), na.rm = TRUE) + c(-buffer, buffer)


# plot
p <- ggplot() +
  # plot landmask
    geom_spatraster(
    data = as.factor(oceanmask)
  ) +
  scale_fill_manual(
    values = c("0" = "white", "1" = "grey10"),
    guide = "none") +
  # plot vessel track
  geom_path(
    data = attraction_seg$vessel,
    aes(x = lon, y = lat, group = id, colour = "Vessel"),
    linewidth = 0.6
  ) +
  geom_point(
    data=attraction_seg$vessel,
    aes(x=lon, y=lat, group = id, colour="Vessel"),
    size=2) +
  # plot animal track
  geom_path(
    data = attraction_seg$animal,
    aes(x = lon, y = lat, group = id, colour = "Animal"),
    linewidth = 0.6
  ) +
  geom_point(
    data=attraction_seg$animal,
    aes(x=lon, y=lat, group = id, colour="Animal"),
    size=2) +
  # plot simulated tracks
  geom_path(
    data=sim_animal_attract,
    aes(x=lon, y=lat, group=id, colour="Simulations"),
    linewidth=0.5, alpha = 0.3) +
  geom_point(
    data=sim_animal_attract,
    aes(x=lon, y=lat, group = id, colour="Simulations"),
    size=1, alpha = 0.3) +
  # set colours
  scale_colour_manual(
    values = c(
      "Animal" = "#0072B2",
      "Vessel" = "#E69F00",
      "Simulations" = "grey"
    ),
    name = NULL
  ) +
  # p-value annotation
  annotate(
    "text",
    x = Inf,
    y = Inf,
    label = p_lab,
    hjust = 1.1,
    vjust = 1.5,
    size = 6
  ) +
  # aesthetics
  coord_sf(
    xlim = xlim_track,
    ylim = ylim_track,
    expand = FALSE
  ) +
  labs(
    x = "Longitude",
    y = "Latitude"
  ) +
  theme_bw()

print(p)
```

Animate plot:

```{r}
p_anim <- p +
  labs(
    x = "Longitude",
    y = "Latitude",
    title = "Time: {frame_along}"
  ) +
  gganimate::transition_reveal(along = time)

gganimate::animate(
  p_anim,
  duration = 6,
  fps = 10,
  renderer = gganimate::gifski_renderer()  # alternative: av_renderer()
)
```

The observed animal trajectory is shown together with the simulated alternative trajectories. If the observed trajectory consistently approaches the vessel more directly than the simulations, this provides evidence supporting attraction behaviour.

# Following analysis

The following test evaluates whether the animal maintained proximity to the vessel longer than expected under random movement once the association had already been established.

In contrast to the attraction analysis, the following test focuses on the persistence of the association rather than on the approach towards the vessel.

The rationale of the test is illustrated below:

1.  the observed trajectory during the association event is extracted;
2.  alternative trajectories are simulated starting from the association onset;
3.  simulated proximity durations are compared against the observed association duration;
4.  if the observed association persists longer than expected under the simulations, the interaction is classified as significant for following behaviour.

## Extract the trajectory segment

For the following analysis, only the observed association event itself is extracted.

```{r}
follow_seg <- extract_event_segments(
  x = association,
  eventID = event_id,
  lead_secs = 0,
  lag_secs = 0
)
```

```{r}
names(follow_seg)
```

```{r}
follow_seg$meta
```


## Run the following test

For the following analysis:

-   `method = "follow"` specifies that we are testing following behaviour;
-   `obs_duration_min` provides the observed association duration used for comparison against the simulations.

```{r}
test_follow <- test_interactions(
  animal = follow_seg$animal,
  vessel = follow_seg$vessel,
  obs_duration_min = follow_seg$meta$duration_min,
  min_dist_m = 1500,
  min_duration_min = 15,
  max_gap_min = 30,
  method = "follow",
  sim_n = 100,
  oceanmask = oceanmask,
  anchor = "start",
  min_locs = 6,
  cores = 2,
  seed = 123,
  return_simdata = TRUE
)
```

## Inspect the output

```{r}
names(test_follow)
```

```{r}
test_follow$result
```

## Interpreting the following test

The following test evaluates whether the observed association duration is longer than expected under the simulated trajectories.

A low p-value indicates that the animal maintained proximity to the vessel more persistently than expected under random movement, supporting following behaviour.

In contrast, non-significant results indicate that the observed association duration could plausibly arise from random movement patterns alone.

## Visualize simulated trajectories

```{r}
sim_animal_follow <- as_track(
  test_follow$sim_pairs,
  lon = animal_lon,
  lat = animal_lat,
  time = animal_time,
  id = matchID
)
```

```{r}

# format p-value for annotation
p_lab <- paste0("p = ", format(round(test_follow$result$p_value, 3), nsmall = 3,
                                 scientific = FALSE, trim = TRUE))

p <- ggplot() +
  # plot vessel track
  geom_path(
    data = follow_seg$vessel,
    aes(x = lon, y = lat, group = id, colour = "Vessel"),
    linewidth = 0.6
  ) +
  geom_point(
    data=follow_seg$vessel,
    aes(x=lon, y=lat, group = id, colour="Vessel"),
    size=2) +
  # plot animal track
  geom_path(
    data = follow_seg$animal,
    aes(x = lon, y = lat, group = id, colour = "Animal"),
    linewidth = 0.6
  ) +
  geom_point(
    data = follow_seg$animal,
    aes(x=lon, y=lat, group = id, colour="Animal"),
    size=2) +
  # plot simulated tracks
  geom_path(
    data=sim_animal_follow,
    aes(x=lon, y=lat, group=id, colour="Simulations"),
    linewidth=0.5, alpha = 0.2) +
  geom_point(
    data=sim_animal_follow,
    aes(x=lon, y=lat, group = id, colour="Simulations"),
    size=1, alpha = 0.2) +
  # set colours
  scale_colour_manual(
    values = c(
      "Animal" = "#0072B2",
      "Vessel" = "#E69F00",
      "Simulations" = "grey"
    ),
    name = NULL
  ) +
  # p-value annotation
  annotate(
    "text",
    x = Inf,
    y = Inf,
    label = p_lab,
    hjust = 1.1,
    vjust = 1.5,
    size = 6
  ) +
  # aesthetics
  coord_sf(expand = FALSE) +
  labs(
    x = "Longitude",
    y = "Latitude"
  ) +
  theme_bw()

print(p)
```


Animate plot:

```{r}
p_anim <- p +
  labs(
    x = "Longitude",
    y = "Latitude",
    title = "Time: {frame_along}"
  ) +
  gganimate::transition_reveal(along = time)

gganimate::animate(
  p_anim,
  duration = 6,
  fps = 10,
  renderer = gganimate::gifski_renderer()  # alternative: av_renderer()
)
```

The observed association trajectory is shown together with the simulated alternative trajectories. If the observed association persists longer than most simulated trajectories, this supports following behaviour.

The classification can be interpreted as follows:

| Attraction | Following | Interpretation | Example dataset |
|---|---|---|---|
| `FALSE` | `FALSE` | The association is not distinguishable from random co-occurrence | #339 |
| `TRUE` | `FALSE` | The animal showed directed movement towards the vessel before the association | #608 |
| `FALSE` | `TRUE` | The animal maintained proximity to the vessel longer than expected | #444 |
| `TRUE` | `TRUE` | The animal both approached the vessel and maintained proximity once associated | #1500 |

This combined interpretation is useful because attraction and following do not necessarily occur together. An animal may approach a vessel but not remain associated with it, or it may remain close to a vessel once encountered without showing clear evidence of directed attraction beforehand.



# Scaling up to multiple animals and vessels

The previous sections used a single animal–vessel pair to illustrate the logic of the method. In real applications, the same workflow is usually repeated across many animals, vessels, trips, and association events.

The general workflow remains the same:

1. match animal and vessel positions;
2. detect associations;
3. extract each association event;
4. run the attraction test;
5. run the following test;
6. store the outputs in a summary table.


## Practical considerations

When scaling up, several additional steps become important:

- use a regional `oceanmask` reused across events;
- store intermediate outputs to avoid losing results if the script stops;
- use `tryCatch()` to handle events where simulations fail;
- set seeds carefully if reproducibility is required;
- avoid storing simulated trajectories for all events unless needed, because this can create very large objects;
- run sensitivity analyses across different distance, duration, and gap thresholds.

For large datasets, the most important change is that the workflow should be treated as a reproducible pipeline rather than as an interactive analysis.

